---
title: "Offshore Indices of Abundance for <br/> Blue Crabs"
author: "Adam A Kemberling"
date: "1/4/2020"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#packages
library(MASS)
library(here)
library(tidyverse)
library(knitr)
library(emmeans)
library(patchwork)
library(gridExtra)
library(broom)
library(statmod)
library(sf)
conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")


#Format ggplot for rest of document
theme_set(theme_bw() + theme(panel.grid.major = element_blank(), 
                             panel.grid.minor = element_blank(), 
                             panel.border = element_rect(fill = NA)))

`%not in%` <- purrr::negate(`%in%`)
```



```{r data}
seamap <- read_csv("~/Dropbox/SEAMAP_2019/data/offshore/seamap_cpue_2019.csv",
                   guess_max = 1e6, 
                   col_types = cols())

#Set up Season Column, using survey titles
seamap[str_detect(seamap$TITLE, "Fall"),"Season"]   <- "Fall"
seamap[str_detect(seamap$TITLE, "Winter"),"Season"] <- "Winter"
seamap[str_detect(seamap$TITLE, "Summer"),"Season"] <- "Summer"
seamap[str_detect(seamap$TITLE, "Spring"),"Season"] <- "Spring"

# Make any setup changes
seamap <- seamap %>% 
  mutate(STAT_ZONE     = factor(STAT_ZONE),
         Crab_presence = ifelse(Sapidus_Catch > 0, 1, 0),
         Survey_Year   = as.numeric(Survey_Year),
         Season        = factor(Season, levels = c("Summer","Fall","Winter")),
         Survey_Month  = lubridate::month(Date),
         Survey_Month  = factor(Survey_Month),
         Survey_design = ifelse(Survey_Year < 2009, "Old Design", "Current Design"),
         Survey_design = ifelse(Survey_Year == 2008 & Season == "Fall", "Current Design", Survey_design),
         Survey_design = factor(Survey_design, levels = c("Old Design", "Current Design")),
         ####  More changes from DE
         Year_f             =  factor(Survey_Year),
         Month_f            =  factor(Survey_Month),
         Depth              =  (Start_Depth + End_Depth)/2,
         Temp_std           =  (Temp_B- mean(Temp_B))/ mean(Temp_B),
         Salinity_std       =  (Salinity_B- mean(Salinity_B))/ mean(Salinity_B),
         Depth_std          =  (Depth- mean(Depth))/ mean(Depth))


# Do all filtering
seamap <- seamap %>% 
  filter(Season            !=  "Winter",
         Survey_Year     %in%  c(2000:2018),
         Spd_kmh           !=  0,                    #these are infinite cpue
         Depth             <=  110)



# Catch only set
crabs <- seamap %>% filter(Sapidus_Catch > 0)


overall_mean <- mean(seamap$CPUE_towspd, na.rm = T)
percent_zeros <- seamap %>% filter(Sapidus_Catch == 0) %>% count()
station_count <- seamap %>% count()
```


# Data Set Summary



 * Total Number of Stations = `r nrow(seamap)`     
 
 * Starting in `r min(as.numeric(as.character(seamap$Survey_Year)))` and continuing through `r max(as.numeric(as.character(seamap$Survey_Year)))`     
 
 * Total Sapidus Caught was `r sum(seamap$Sapidus_Catch, na.rm = T)`     
 
 * Overall Frequency of Occurrence was `r mean(seamap$Crab_presence, na.rm = TRUE) * 100`     
 
```{r}
crabs %>% 
  mutate(STAT_ZONE = fct_rev(STAT_ZONE)) %>% 
  ggplot(aes(STAT_ZONE, Sapidus_Catch)) + 
  geom_boxplot(aes(group = STAT_ZONE)) + 
  labs(title    = "Blue Crab Catch by Stat Zone", 
       subtitle = "Stat zones ordered West to East - 2000 to 2018", 
       x        = "Statistical Zone",
       y        = "C. sapidus Single Station Catch") +
  facet_wrap(~Survey_design)
```
 
 
 

# Factor Summaries

## Yearly Summaries {.tabset .tabset-fade .tabset-pills}

**Overall Mean Catch Crabs/Hectare:** `r overall_mean`   

**Stations with no catch:** `r percent_zeros` / `r station_count` or `r (percent_zeros / station_count) * 100`      

```{r}
year_sums <- seamap %>% 
  group_by(Survey_Year) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(CPUE_towspd)),
            `Mean Catch`      = mean(CPUE_towspd, na.rm = T),
            `SD`              = sd(CPUE_towspd, na.rm = T),
            `Var`             = var(CPUE_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(CPUE_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(Crab_presence, na.rm = T),
            `% Occurrence SE` = sd(Crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
  ungroup() %>% 
    mutate(Survey_Year        = factor(Survey_Year),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )


year_sums %>% select(Survey_Year, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, `Overdispersed`) %>% knitr::kable()
#knitr::kable(year_sums)
```

***

### Yearly Catch Plot


```{r}

year_plot_1 <- year_sums %>% 
  ggplot() +
    geom_col(aes(Survey_Year, `Mean Catch`)) +
    geom_errorbar(aes(x = Survey_Year, 
                      ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
  labs(x = "Survey Year",
       y = "Mean Catch (C. sapidus / Hectare)") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))


year_plot_1
#ggsave(plot =  year_plot_1, here("2019_edits/figures", "yearly_catch.png"), dpi = 600)

```

***

### Yearly Frequency of Occurrence

```{r}

year_plot_2 <- year_sums %>% 
  ggplot() +
    geom_col(aes(Survey_Year, `% Occurrence`)) +
    geom_errorbar(aes(x = Survey_Year, 
                      ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  size = .25) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = "Survey Year",
       y = "Percent Occurrence") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

year_plot_2
#ggsave(plot = year_plot_2, here("2019_edits/figures", "yearly_occurrence.png"), dpi = 600)

```

***


### Yearly Side-by

```{r}
year_plot_1 + year_plot_2
```


***

## Stat Zone Summaries {.tabset .tabset-fade .tabset-pills}


```{r}
#Assign Stat Zone from NMFS grid
szones <- read_sf("~/Dropbox/SEAMAP_2019/data/gis_files/NMFS_Fishing_Grids_GOM_2013.shp", 
                  crs = 4326)

# Add a better StatZone ID by spatial overlap
seamap <- seamap %>% 
  st_as_sf(coords = c("Start_Long", "Start_Lat"), crs = 4326, remove = FALSE) %>% 
  st_join(szones["StatZone"]) %>% 
  st_set_geometry(NULL) %>% 
  as.data.frame() %>% 
  dplyr::select(-STAT_ZONE) %>%  #Remove default/incorrect one
  mutate(StatZone = factor(StatZone))



szone_summary <- seamap %>% 
  filter(is.na(StatZone)      == FALSE) %>% 
  group_by(StatZone) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(CPUE_towspd)),
            `Mean Catch`      = mean(CPUE_towspd, na.rm = T),
            `SD`              = sd(CPUE_towspd, na.rm = T),
            `Var`             = var(CPUE_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(CPUE_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(Crab_presence, na.rm = T),
            `% Occurrence SE` = sd(Crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
    ungroup() %>% 
    mutate(StatZone           = factor(StatZone),
           StatZone           = fct_rev(StatZone),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )

szone_summary %>% 
  select(StatZone, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% 
  knitr::kable()

```

***

### Stat Zone Catch Plot

```{r}

szones_plot_1 <- szone_summary %>% 
  ggplot() +
    geom_col(aes(StatZone, `Mean Catch`)) +
    geom_errorbar(aes(x = StatZone, 
                      ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
  geom_vline(xintercept = 11.5, color = "gray", lty = 2) +
  geom_vline(xintercept = 4.5, color = "gray", lty = 2) +
  #annotate(geom = "text", x = 7, y = 2.0, label = "LA + MS") +
  #annotate(geom = "text", x = 16, y = 2.0, label = "East of Mobile Bay") +
  annotate(geom = "text", x = 16.5, y = -0.1, label = "East of Mobile Bay") +
  annotate(geom = "text", x = 8, y = -0.1, label = "LA & MS") +
  annotate(geom = "text", x = 2.5, y = -0.1, label = "Texas") +
    labs(x= "Statistical Zone",
         y = "Mean Catch (C. sapidus / Hectare)")


szones_plot_1 
#ggsave(plot = szones_plot_1, here("2019_edits/figures", "statzones_catch.png"), dpi = 600)

```


***

### Stat Zone Frequency of Occurrence

```{r}

szones_plot_2 <- szone_summary %>% 
  ggplot() +
    geom_col(aes(StatZone, `% Occurrence`)) +
    geom_errorbar(aes(x = StatZone, 
                      ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  size = .25) +
  geom_vline(xintercept = 11.5, color = "gray", lty = 2) +
  geom_vline(xintercept = 4.5, color = "gray", lty = 2) +
  annotate(geom = "text", x = 16.5, y = 0.75, label = "East of Mobile Bay") +
  annotate(geom = "text", x = 8, y = 0.75, label = "LA & MS") +
  annotate(geom = "text", x = 2.5, y = 0.75, label = "Texas") +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = "Statistical Zone",
       y = "Percent Occurrence")

szones_plot_2
#ggsave(plot = szones_plot_2, here("2019_edits/figures", "statzones_occurrence.png"), dpi = 600)
```


***

###  Stat Zone Side-by

```{r}
szones_plot_1 + szones_plot_2
```



***


## Region Summary {.tabset .tabset-fade .tabset-pills}


```{r}
seamap <- seamap %>% 
  mutate(
    # region = ifelse(Start_Lat < 28.387595 & Start_Long < -96, "South Texas", "TX/LA"),
    #      region = ifelse(Start_Long > -89.2, "MS Bight", region),
    #      region = ifelse(Start_Long > -87.5, "FL", region),
    #      region = factor(region)
    region = ifelse(StatZone %in% seq("21", "18"), "Texas", "Louisiana"),
    region = ifelse(StatZone %in% seq("12", "10"), "MS Bight", region),
    region = ifelse(StatZone %in% seq("1", "9"), "Florida", region),
    region = factor(region, levels = c("Texas", "Louisiana", "MS Bight", "Florida"))
    )

#Output seamap analysis file
write_csv(seamap, 
          "~/Dropbox/SEAMAP_2019/data/offshore/seamap_cpue_2019_analysis.csv", 
          col_names = TRUE)

region_summary <- seamap %>% 
  filter(is.na(region)      == FALSE) %>% 
  group_by(region) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(CPUE_towspd)),
            `Mean Catch`      = mean(CPUE_towspd, na.rm = T),
            `SD`              = sd(CPUE_towspd, na.rm = T),
            `Var`             = var(CPUE_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(CPUE_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(Crab_presence, na.rm = T),
            `% Occurrence SE` = sd(Crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
    ungroup() %>% 
    mutate(region             = factor(region, 
                                       levels = c("Texas", "Louisiana", "MS Bight", "Florida")),
           #region             = fct_rev(region),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
    )

region_summary %>% 
  select(region, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% 
  knitr::kable()

```

***

### Regional Catch Plot


```{r}

region_plot_1 <- region_summary %>% 
  ggplot() +
    geom_col(aes(region, `Mean Catch`)) +
    geom_errorbar(aes(x = region, 
                      ymin = catch_ymin,
                      ymax = catch_ymax),
                  size = .25) +
    labs(x= "Sampling Region",
         y = "Mean Catch (C. sapidus / Hectare)")


region_plot_1 
#ggsave(plot = region_plot_1, here("2019_edits/figures", "regions_catch.png"), dpi = 600)
```

***

### Regional Frequency of Occurrence

```{r}

region_plot_2 <- region_summary %>% 
  ggplot() +
    geom_col(aes(region, `% Occurrence`)) +
    geom_errorbar(aes(x = region, 
                      ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  size = .25) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = "Sampling Region",
       y = "Percent Occurrence")

region_plot_2
#ggsave(plot = region_plot_2, here("2019_edits/figures", "regions_occurrence.png"), dpi = 600)

```

***

### Regional Side-by

```{r}
region_plot_1 + region_plot_2
```


***


## Region - Season {.tabset .tabset-fade .tabset-pills}

```{r}
rs_summary <- seamap %>% 
  filter(is.na(region)      == FALSE) %>% 
  group_by(region, Season) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(CPUE_towspd)),
            `Mean Catch`      = mean(CPUE_towspd, na.rm = T),
            `SD`              = sd(CPUE_towspd, na.rm = T),
            `Var`             = var(CPUE_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(CPUE_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(Crab_presence, na.rm = T),
            `% Occurrence SE` = sd(Crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
    ungroup() %>% 
    mutate(Season             = factor(Season, levels = c("Summer", "Fall")),
           region             = factor(region, 
                                       levels = c("Texas", "Louisiana", "MS Bight", "Florida")),
           #region             = fct_rev(region),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
    )

rs_summary %>% 
  select(region, Season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% 
  knitr::kable()
```


***

### Region - Season Catch

```{r}
rs_plot_1 <- rs_summary %>% 
  ggplot(aes(region, `Mean Catch`, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  position = "dodge", 
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = "Survey Year",
       y = "Mean Catch (C. sapidus / Hectare)") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA),
    axis.text.x = element_text(angle = 45, vjust = 0.5)
    )


rs_plot_1 
```


***

### Region - Season FOO

```{r}

rs_plot_2 <- rs_summary %>% 
  ggplot(aes(region, `% Occurrence`, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = "Survey Year",
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA),
    axis.text.x = element_text(angle = 45, vjust = 0.5)
    )

rs_plot_2

```


***

### Region - Season Side-by


```{r}

rs_plot_1 + theme(legend.position = "none") + rs_plot_2

```

***


## Seasonal Summary {.tabset .tabset-fade .tabset-pills}



```{r}

season_summary <- seamap %>% 
  group_by(Survey_Year, Season) %>% 
  summarise(Stations          = n(),
            `Missing Obs`     = sum(is.na(CPUE_towspd)),
            `Mean Catch`      = mean(CPUE_towspd, na.rm = T),
            `SD`              = sd(CPUE_towspd, na.rm = T),
            `Var`             = var(CPUE_towspd),
            `Overdispersed`   = ifelse(Var > `Mean Catch`, "yes", "no"),
            `Catch SE`        = sd(CPUE_towspd, na.rm = T) / sqrt(Stations - `Missing Obs`),
            `% Occurrence`    = mean(Crab_presence, na.rm = T),
            `% Occurrence SE` = sd(Crab_presence, na.rm = T) / sqrt(Stations - `Missing Obs`)) %>% 
    ungroup() %>% 
    mutate(Survey_Year        = factor(Survey_Year),
           # StatZone         = factor(StatZone),
           # StatZone         = fct_rev(StatZone),
           catch_ymin         = `Mean Catch` - (2 * `Catch SE`),
           catch_ymax         = `Mean Catch` + (2 * `Catch SE`),
           catch_ymin         = ifelse(catch_ymin < 0, 0, catch_ymin),
           occurrence_ymin    = `% Occurrence` - (2 * `% Occurrence SE`),
           occurrence_ymax    = `% Occurrence` + (2 * `% Occurrence SE`),
           occurrence_ymin    = ifelse(occurrence_ymin < 0, 0, occurrence_ymin)
         )

season_summary %>% select(Survey_Year, Season, Stations, `Missing Obs`, `% Occurrence`, `Mean Catch`, SD, Overdispersed) %>% knitr::kable()
#knitr::kable(season_summary)
```

***

### Seasonal Catch Plot

```{r}

season_plot_1 <- season_summary %>% 
  ggplot(aes(Survey_Year, `Mean Catch`, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = catch_ymin,
                      ymax = catch_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  labs(x = "Survey Year",
       y = "Mean Catch (C. sapidus / Hectare)") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA),
    axis.text.x = element_text(angle = 45, vjust = 0.5)
    )


season_plot_1 
#ggsave(plot =  year_plot_1, here("2019_edits/figures", "yearly_catch.png"), dpi = 600)


```

***

### Seasonal Frequency of Occurrence

```{r}

season_plot_2 <- season_summary %>% 
  ggplot(aes(Survey_Year, `% Occurrence`, fill = Season)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = occurrence_ymin,
                      ymax = occurrence_ymax),
                  position = "dodge",
                  size = .25) +
  scale_fill_manual(name = "", values = setNames(c("gray35", "gray50"), c("Summer", "Fall"))) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = "Survey Year",
       y = "Percent Occurrence") + theme(
    legend.position = c(1, .95),
    legend.justification = c("right", "top"),
    legend.box.just = "right",
    legend.margin = margin(0, 6, 2, 2),
    legend.background = element_rect(fill = NA),
    axis.text.x = element_text(angle = 45, vjust = 0.5)
    )

season_plot_2



```

***

### Seasonal Side-by

```{r}
season_plot_1 + theme(legend.position = "none") + season_plot_2
```


***


# Crab/hectare Models {.tabset .tabset-pills}


## Model A. 2000-2018 w/ interaction

This model uses crabs per hectare as the response over the period of 18 years (2000-2018). Data is limited to statzones that were consistently sampled over the time period (i.e. West of Mobile Bay).

Model is a generalized linear model with a negative binomial error distribution. Blue crab catch is predicted as the sum of linear predictors for the year and season when the station was sampled in addition to the region in which it took place, with an interaction between survey year and region. An offset of area towed in hectares was used to predict catch as a rate and account for differences in sampling effort at different stations.

```{r}
seamap <- seamap %>% rename_all(tolower)
seamap_west <- seamap %>% 
  filter(statzone != 12,
         statzone %in% 11:21) %>% 
  mutate(statzone = factor(statzone))

mod1a <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            year_f * region + 
            offset(log(area_hectares)),
            data = seamap_west)

#summary(mod1a)
```

### Interaction Plot

Looking at an interaction plot indicates that the impact of each region is not consistent across years and that there is an important interaction in the data.

```{r}
emmeans::emmip(mod1a, region ~ year_f) +
  labs(title = "Interaction Plot: year * region",
       caption = "criss-crossing of estimates is subjective indication of interaction",
       x = NULL) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

### Model Summary

Understanding p-values in glm model summary. Using the default settings in R sets the contrast levels for factors such that all factors are contrasted against the first/lowest factor level.

For example in our model year is modeled as a factor, which adjusts the intercept for the model for each. The first year is 2000, so all other years are contrasted against that year's mean estimate. One quick way to confirm what factor level is the contrast is y checking which levels are absent from the summary table `summary(my_model)`

Contrasts:   
year = 2000   
region = texas   
season = summer   

P-values on factor levels in these tables indicate significant deviation from the contrasts. Post-hoc tests are necessary to compare individual levels against one another.

```{r}
tidy(mod1a) %>% 
  mutate(`signif.` = ifelse(p.value <= .05, "*", "")) %>% 
  kable()
```

### Fit Diagnostics - Deviance



```{r}
broom::glance(mod1a) %>% kable()

#Percent deviance explained
perc_dev <- (mod1a$null.deviance - mod1a$deviance) / mod1a$null.deviance
```

Goodness of fit for generalized linear models is described using deviance. A glm will report 2 deviance measures: 

`null.deviance` is the residual variation remaining from the null model. The null model here is the grand mean of blue crab catch per hectare of all observations

`deviance` is the residual variation remaining from the fitted model. Numbers closer to the null.deviance indicate that the model did not perform much better than the null model.

`% deviance explained` is calculated as $\frac{null.deviance - deviance}{null.deviance} * 100$ and expresses the improvement of the fitted model over the null in % deviance explained.

Percent Deviance for model A. = `r perc_dev`

#### CPUE Timeline

Because we used area towed in the model as an offset predictions from our model can be interprated as crabs per hectare.

```{r}

#Estimated Marginal Means with an offset
#parameter in the call of offset = 0:
mod1a.dat <- as.data.frame(emmeans(mod1a, 
                                  specs = "year_f", 
                                  by    = "region",
                                  type  = "response", 
                                  offset = 0))
mod1a.dat$CV <- mod1a.dat$SE/mod1a.dat$response







#make scale_x_discrete breaks b/c ggplot is dumb
ystart <- min(as.numeric(seamap_west$survey_year))
yend   <- max(as.numeric(seamap_west$survey_year))
ynames <- lubridate::year(seq.Date(as.Date(paste(ystart), "%Y"), 
                   as.Date(paste(yend), "%Y"), 
                   by = "year"))



#ggplot of standardized index of abundance
p <- mod1a.dat %>% 
  mutate(year_n = as.numeric(as.character(year_f))) %>% 
  ggplot(aes(year_n, response)) +
      geom_ribbon(aes(year_n, ymin = asymp.LCL, ymax = asymp.UCL), fill = "gray80") +
      geom_line(aes(year_n, response, group = 1)) +
      geom_point(aes(year_n, response)) +
      geom_hline(aes(yintercept = mean(seamap_west$cpue_towspd)), color = "darkred", linetype = 2) +
      scale_x_discrete(limits = c(as.numeric(ystart:yend)), labels = ynames) +
      labs(
        x = NULL, 
        y = "Crabs per Hectare") +
      theme(axis.text.x=element_text(angle=45, hjust = 1, vjust=1), legend.position = "bottom") +
      facet_wrap(~region)
p

```

### Observed vs. Predicted, all model factors

Model prediction (black) & observed mean values (blue), both standardized by overall mean

```{r}

#Estimated Marginal Means with an offset
#parameter in the call of offset = 0:
mod1a.dat <- as.data.frame(emmeans(mod1a, 
                                  specs = "year_f", 
                                  by    = c("region", "season"),
                                  type  = "response", 
                                  offset = 0))
mod1a.dat$CV <- mod1a.dat$SE/mod1a.dat$response


#Observed data
mean_obs <- mean(seamap_west$cpue_towspd)
obs_timeline <- seamap_west %>% 
  group_by(survey_year, region, season) %>% 
  summarise(Nobs = n(),
            PercentPos = mean(crab_presence),
            Mean_CPUE = mean(cpue_towspd)) %>% 
  mutate(Index_Obs = Mean_CPUE/mean_obs)





#ggplot of standardized index of abundance
p <- mod1a.dat %>% 
  mutate(year_n = as.numeric(as.character(year_f))) %>% 
  ggplot(aes(year_n, response)) +
      #Fitted Data
      geom_ribbon(aes(year_n, ymin = asymp.LCL, ymax = asymp.UCL), fill = "gray80") +
      geom_line(aes(year_n, response, group = 1)) +
      geom_point(aes(year_n, response)) +
      # Observed Data
      geom_line(data = obs_timeline, 
              aes(survey_year, Mean_CPUE, group = region), 
              col = "royalblue", linetype = 2) +
      geom_point(data = obs_timeline, 
                 aes(survey_year, Mean_CPUE, group = region), col = "royalblue") +
      #Theme changes
      scale_x_discrete(limits = c(as.numeric(ystart:yend)), labels = ynames) +
      labs(
        x = NULL, 
        y = "Crabs per Hectare",
        title = "Model Estimates and Yearly observed means") +
      theme(axis.text.x=element_text(angle=45, hjust = 1, vjust=1), legend.position = "bottom") +
      facet_grid(season~region)
p

```



## Model B. 2009-2018 

This model uses crabs per hectare as the response over the period of 9 years (2009-2018). Data selection and sampling effort was consistent over this period. Additionally the survey expanded to the Eastern Gulf of Mexico so stat zones in that region are included as well.

Model is a generalized linear model with a negative binomial error distribution. Blue crab catch is predicted as the sum of linear predictors for the year and season when the station was sampled in addition to the region in which it took place, with an interaction between survey year and region. An offset of area towed in hectares was used to predict catch as a rate and account for differences in sampling effort at different stations.


```{r}
seamap_east <- seamap %>% 
  filter(statzone != 12,
         statzone != 1,
         survey_year %in% 2009:2018) %>% 
  mutate(statzone = factor(statzone))

modb <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            year_f * region + 
            offset(log(area_hectares)),
            data = seamap_east)

#summary(modb)
```

### Interaction Plot

Looking at an interaction plot indicates that the impact of each region is not consistent across years and that there is an important interaction in the data. The sharp peaks for Florida's linear predictor reflect years when there was no catch.

```{r}
emmeans::emmip(modb, region ~ year_f) +
  labs(title = "Interaction Plot: year * region",
       caption = "criss-crossing of estimates is subjective indication of interaction",
       x = NULL) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```

### Model Summary


```{r}
tidy(modb) %>% 
  mutate(`signif.` = ifelse(p.value <= .05, "*", "")) %>% 
  kable()
```

### Fit Diagnostics - Deviance


```{r}
broom::glance(modb) %>% kable()

#Percent deviance explained
perc_dev <- (modb$null.deviance - modb$deviance) / modb$null.deviance
```

Goodness of fit for generalized linear models is described using deviance. A glm will report 2 deviance measures: 

`null.deviance` is the residual variation remaining from the null model. The null model here is the grand mean of blue crab catch per hectare of all observations

`deviance` is the residual variation remaining from the fitted model. Numbers closer to the null.deviance indicate that the model did not perform much better than the null model.

`% deviance explained` is calculated as $\frac{null.deviance - deviance}{null.deviance} * 100$ and expresses the improvement of the fitted model over the null in % deviance explained.

Percent Deviance for model B. = `r perc_dev`

#### CPUE Timeline

Because we used area towed in the model as an offset predictions from our model can be interprated as crabs per hectare.

```{r}

#Estimated Marginal Means with an offset
#parameter in the call of offset = 0:
modb.dat <- as.data.frame(emmeans(modb, 
                                  specs = "year_f", 
                                  by    = "region",
                                  type  = "response", 
                                  offset = 0))
modb.dat$CV <- modb.dat$SE/modb.dat$response







#make scale_x_discrete breaks b/c ggplot is dumb
ystart <- min(as.numeric(seamap_east$survey_year))
yend   <- max(as.numeric(seamap_east$survey_year))
ynames <- lubridate::year(seq.Date(as.Date(paste(ystart), "%Y"), 
                   as.Date(paste(yend), "%Y"), 
                   by = "year"))



#ggplot of standardized index of abundance
p <- modb.dat %>% 
  mutate(year_n = as.numeric(as.character(year_f))) %>% 
  ggplot(aes(year_n, response)) +
      geom_ribbon(aes(year_n, ymin = asymp.LCL, ymax = asymp.UCL), fill = "gray80") +
      geom_line(aes(year_n, response, group = 1)) +
      geom_point(aes(year_n, response)) +
      geom_hline(aes(yintercept = mean(seamap_east$cpue_towspd)), color = "darkred", linetype = 2) +
      scale_x_discrete(limits = c(as.numeric(ystart:yend)), labels = ynames) +
      labs(
        x = NULL, 
        y = "Crabs per Hectare") +
      theme(axis.text.x=element_text(angle=45, hjust = 1, vjust=1), legend.position = "bottom") +
      facet_wrap(~region)
p

```

### Observed vs. Predicted, all model factors

Model prediction (black) & observed mean values (blue), both standardized by overall mean

```{r}

#Estimated Marginal Means with an offset
#parameter in the call of offset = 0:
modb.dat <- as.data.frame(emmeans(modb, 
                                  specs = "year_f", 
                                  by    = c("region", "season"),
                                  type  = "response", 
                                  offset = 0))
modb.dat$CV <- modb.dat$SE/modb.dat$response

#Drop Infinite credibility intervals
modb.dat <- modb.dat %>% 
  mutate(asymp.UCL = ifelse(asymp.UCL == Inf, NA, asymp.UCL))


#Observed data
mean_obs <- mean(seamap_east$cpue_towspd)
obs_timeline <- seamap_east %>% 
  group_by(survey_year, region, season) %>% 
  summarise(Nobs = n(),
            PercentPos = mean(crab_presence),
            Mean_CPUE = mean(cpue_towspd)) %>% 
  mutate(Index_Obs = Mean_CPUE/mean_obs)





#ggplot of standardized index of abundance
p <- modb.dat %>% 
  mutate(year_n = as.numeric(as.character(year_f))) %>% 
  ggplot(aes(year_n, response)) +
      #Fitted Data
      geom_ribbon(aes(year_n, ymin = asymp.LCL, ymax = asymp.UCL), fill = "gray80") +
      geom_line(aes(year_n, response, group = 1)) +
      geom_point(aes(year_n, response)) +
      # Observed Data
      geom_line(data = obs_timeline, 
              aes(survey_year, Mean_CPUE, group = region), 
              col = "royalblue", linetype = 2) +
      geom_point(data = obs_timeline, 
                 aes(survey_year, Mean_CPUE, group = region), col = "royalblue") +
      #Theme changes
      scale_x_discrete(limits = c(as.numeric(ystart:yend)), labels = ynames) +
      labs(
        x = NULL, 
        y = "Crabs per Hectare",
        title = "Model Estimates and Yearly observed means") +
      theme(axis.text.x=element_text(angle=45, hjust = 1, vjust=1), legend.position = "bottom") +
      facet_grid(season~region)
p

```
