---
title: "NOAA Tagging Final Report"
author: "Adam A Kemberling"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    mathjax: null
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

#Packages
library(tidyverse)
library(here)
library(stargazer)
library(sf)
library(rnaturalearth)
library(gridExtra)
library(circular)


#Data
tagging <- read_csv(here("data", "noaa_report", "tags_clean_final.csv"), 
                    col_types = cols())
recaps <- read_csv(here("data", "noaa_report", "tags_recaps_clean_final.csv"), 
                   col_types = cols())
both <- read_csv(here("data", "noaa_report", "tags_recaps_clean_final.csv"), 
                 col_types = cols())
usa_poly <- ne_states("united states of america") %>% st_as_sf() %>% filter(region == "South")
mexico_poly <- ne_states("mexico") %>% st_as_sf()
```

# Gulf Scale


##  Summary Statistics

o   Total numbers tagged recaptured

o   Time at liberty

o   Distance travelled (water only)

o   Travel rates

```{r}
both %>% 
  group_by(tag_state) %>% 
  summarise(`Average Time at Liberty` = mean(time_at_large, na.rm = T),
            `Average Distance Travelled (by water*)` = mean())
```


##   Tagging State and Recapture State (Table or alluvial Plot)

 
# State Scale

##   Summary Statistics

o   Total numbers tagged recaptured

o   Time at liberty

o   Distance travelled (water only)

o   Travel rates


```{r}

```

##   Tagging State and Recapture State (Table or alluvial Plot)

```{r}

```
 

# Basin Scale

##   Summary Statistics

o   Total numbers tagged when (maybe a figure with calendar month and numbers tagged

o   Maps of tagging and recaptures

```{r}

```


##   Directions travelled (exclude crabs that travelled < 5km for this)

o   Rao’s test for each

o   Crabs recaptured within home estuary

o   Crabs captured somewhere else

##  Travel rates (Maybe exclude crabs that have long times at liberty? Start with 1 month)

o   What patterns stand out: mean and variance etc. by season

```{r}

```

 
# Crabs outside of Estuaries

##   Map of just them

```{r}

```


##   Any recaptures

```{r}

```

